### 2022-02-11 article/article.tex
version with 6 pages, Abstr/Intro/Concl TODO

% \loop\iftrue%
% \errmessage{This article is copyrighted by ___ and should not be TeXed}%
% \repeat%

\documentclass{ifacconf}  % Comment this line out
\usepackage{natbib}        % required for bibliography

% \documentclass[a4paper, 10 pt, conference]{ieeeconf}  % Comment this line out
% \IEEEoverridecommandlockouts%
\pdfminorversion=4%

\input{preamble}
\input{math}
\input{symbols}

% \usepackage{flushend}% to equalize last page %%% use when text complete

\draft% show todos in red
% \final% give error if there is todos
% \comments% blank right column to make comments in pen & paper

% \newcommand{\id}{}
% \newcommand{\price}{31.00}
% \pubid{\begin{minipage}{\textwidth}\ \\[50pt]\centering
%   \copyright{} 20XX \_\_\_\_.\\
%   Personal use is permitted, but republication/redistribution requires permission.\\
%   % See https://www.ieee.org/publications/rights/index.html for more information.% if IEEE
% \end{minipage}
% }
% \makeglossaries

\begin{document}

\begin{frontmatter}
\title{\LARGE \bf
  Expectation-Maximization based defense mechanism for distributed Model Predictive Control
}

\author[First]{Rafael Accácio Nogueira}
\author[First]{Romain Bourdais}
\author[First]{Simon Leglaive}
\author[First]{Hervé Guéguen}
\address[First]{IETR-CentraleSupélec, 35510 Cesson-Sévigné, Ille-et-Vilaine, France\\
{\tt\small \{rafael-accacio.nogueira, romain.bourdais, simon.leglaive, herve.gueguen\}
@centralesupelec.fr}}

% \thanks{\centering The authors are with  \newline {\tt\small
% }\newline

% % <-this % stops a space
% \thanks{\centering The authors are with IETR-CentraleSupélec, 35510\newline Cesson-Sévigné, Ille-et-Vilaine, France \newline {\tt\small
% \{rafael-accacio.nogueira, romain.bourdais, herve.gueguen\}
% @centralesupelec.fr}\newline
% }%
% }

% \maketitle
% \IEEEpeerreviewmaketitle%
% \thispagestyle{empty}
% \pagestyle{empty}

\begin{abstract}% Abstract of not more than 250 words.
  \todo[write abstract]{
  In distributed predictive control structures, communication among agents is required to achieve a consensus and approach an optimal global behavior. Such negotiation mechanisms are sensitive to attacks on these exchanges.
  This paper proposes a monitoring scheme based on Expectation Maximization to detect and mitigate the effects of attacks in a resource allocation framework.
  The performance is assessed through academical example of the temperature control of multiple rooms under input power constraints.
}
\end{abstract}

\begin{keyword}
  Model predictive and optimization-based control,
  Distributed control and estimation,
Large scale optimization problems
\end{keyword}

\end{frontmatter}

\section{INTRODUCTION}
The information collected, processed, and transmitted within \cps{} impacts physical systems whose proper functioning is essential for society, like energy networks, industrial processes, water production plants, etc..
Attacks in those critical infrastructures, such as Stuxnet, increased awareness for cyber-security of \cps{}.

Given the complexity, scale and even geographic location of these structures, it is necessary to adopt control structures that are decentralized rather than \todo{monolithic}.
For these systems, a set of controllers are designed to ensure the safe and efficient operation of the facilities collectively.

This operation is done cooperatively, where the controllers exchange information with each other.
This approach can be used, for example, for the voltage regulation of electrical networks, and the production-consumption balance in these same networks.
Multi-agent systems~\citep{KantamneniEtAl2015} or \dmpc{}~\citep{MaestreEtAl2014} are standard tools to design these controllers.

A few articles have focused on safety in \dmpc{} frameworks, presenting vulnerabilities of different frameworks, such as Lagrange-based decomposition~\citep{VelardeEtAl2017}, Jacobi-Gauss decomposition~\citep{ChanfreutEtAl2018} and primal decomposition~\citep{NogueiraEtAl2021}.

The authors in these works highlight the complexity of the problems. A modification of a local goal, the hacking of communication, or the failure of one of the agents are events that disrupt global behavior.
If one of these agents is attacked, it can result in the violent destruction of the corrupted element (perhaps the whole system), or more subtly, it can lead to deviant behavior that is more difficult to detect.
% Usually the attacks aim to de-optimize the global operation by benefiting/prejudicing some specific group of controllers, or to lead the system towards a state of malfunction/instability~\todo[use other]{}.

Some strategies are presented to mitigate the effects of such event.
For instance dismissing extreme values as in~\cite{VelardeEtAl2017}, or
using secure scenarios based on reliable historical data~\cite{MaestreEtAl2021}.

Our previous work~\citep{NogueiraEtAl2021} proposed a safe algorithm based on data reconstruction to mitigate the effects of a \fdi{} in the communication between a coordinator and local constraint-free agents, which were bound by global equality constraints.

In this paper, we propose an extension of our approach by including local constraints and changing the global constraints.
This extension profoundly changes the complexity, as the exchanges between the
agents and the coordinator are no longer characterized by an affine functions
but by \pwa{} functions.
This leads not only to a combinatorial explosion but also to a parametric identification challenge.
To overcome this problem, we propose using a learning method based on the \EM{} algorithm~\citep{DempsterEtAl1977}, which allows us to estimate the corruption
mechanism and correct it if necessary.

\todo{
This paper is organized as follows.
First, in Section~\ref{sec:PS}, we introduce the primal decomposition-based DMPC and its vulnerabilities.
Then in Section~\ref{sec:TSM} we study the local problems and propose a detection and mitigation mechanism based on the Expectation Maximization algorithm.
Moreover, in Section~\ref{sec:App_dMPC}, we illustrate the algorithm with academic simulation and we assess its performance.
Finally, in Section~\ref{sec:CC}, we conclude, and we give an outlook of future works.
}

\emph{Notation:}
In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and $\norm{\vec{v}}_{Y}$ represent the $\ell_{2}$ norm, the Frobenius norm and the weighted norm $\norm{Y^{\frac{1}{2}}\vec{v}}$.
$\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto $\set{T}$.
% $\card{x}$ is the number of elements in $x$.
$\kron{}{}$ ($\hadamard{}{}$) is the Kronecker (Hadamard) product.
$\vectorize{A}$ vectorizes matrix $A$.
${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ is equal to ${n\mathbin{:}1\mathbin{:}j}$.
$[A,B]$ ($[A;B]$) is horizontal (vertical) concatenation of matrices.
${\0_{m}=0_{m\times 1}}$ (${\1_{m}=1_{m\times 1}}$), where
$0_{m\times n}$ ($1_{m\times n}$) is a $0$ ($1$) filled \mbox{$m$-by-$n$} matrix.
$\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
$\preceq$ is the component-wise inequality for vectors and matrix inequality for matrices in ${\symmetric^{n}=\setbuild{X:\R^{n\times n}}{X\T=X}}$.
${\defpos^{n}=\setbuild{X:\symmetric^{n}}{X\succ 0}}$, ${\semidefpos^{n}=\setbuild{X:\symmetric^{n}}{X\succeq 0}}$.
$I_{c}$ is the identity matrix of size $c$.
$A^{\dagger}={{(A^{T}A)}^{-1}A^{T}}$.
$\determinant{A}$ is the determinant of $A$.
A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
$\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
$\expectation{\random{x}}$ is the expected value of $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
% $\pdf{\cdot}$ is a probability density function.
${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.

\section{Problem Statement}\label{sec:PS}
In this section we present an algorithm to decompose a \mpc{} problem based on the primal decomposition \citep{BoydEtAl2015}, and a vulnerability of the decomposition that can be exploited by a malicious agent.

\subsection{Model Predictive Control}\label{ssec:MPC}
Consider a system composed of discrete-time linear time-invariant agents ${i\in\set{M}=\{1\mathbin{:}\nsubsystems\}}$, modeled by
\begin{equation}
\begin{matrix}
  \label{eq:systems}
\vec{x}_{i}[k+1]&=&A_{i}\xik &+& B_{i}\uik\\
\vec{y}_{i}[k+1]&=&C_{i}\xik &&
\end{matrix},
\end{equation}
with $\vec{x}_{i}[k]:\R^{n_{x}}$, $\vec{u}_{i}[k]:\R^{n_{u}}$ and $\vec{y}_{i}[k]:\R^{n_{y}}$.

Each agent's input $\uik$ is constrained by
\begin{equation}
  \label{eq:u_local_constraints}
  \0_{n_{u}}\preceq\uik\preceq\vec{u}_{\max},
\end{equation}
with ${\umax:\R^{n_{u}}}$.
The agents are coupled by global input constraints with weighting matrices \todo{${\Gamma_{i}:\semidefpos^{n_{u}}}$}:
\begin{equation}
  \label{eq:global_constraint}
  \sum_{i\in\set{M}}\Gamma_{i}\uik\preceq\umax.
\end{equation}

The overall system is controlled by a \mpc{}, which
computes the optimal input for a finite prediction horizon ${\set{H}=\{1\mathbin{:}\predhorz\}}$ by solving the following problem:
\begin{equation}\label{eq:GOP}
\resizebox{.88\columnwidth}{!}{$
\begin{matrix}
\underset{\useq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits_{i\in\set{M}} \overbrace{\sum_{j\in\set{H}}\norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
\mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:u_local_constraints}\ \mathrm{and}\ \eqref{eq:global_constraint}
\left\}\small
\begin{aligned}
  &\forall i\in\set{M}\\
  &\forall j\in\set{H}
\end{aligned}\right.,

\end{matrix}
  $}
% \label{eq:dmpcModLOP}
\end{equation}
where ${Q_{i}:\semidefpos}$, ${R_{i}:\defpos}$, and $\vik$ is a control objective.
$\optglobobj$ denotes the optimal value of the objective function for the the optimal control sequences ${\optuiseq}$.
The problem in~\eqref{eq:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in their respective subsystem $i$, following a \rhs{}.

Solving~\eqref{eq:GOP} at each time $k$ can be quite challenging, since the prediction horizon and the number of agents can be large making the process computationally intensive. Another issue that can arise is the fact that the complete information \todo[verify]{(${\set{I}_{i}=\{A_{i},B_{i},C_{i},\vec{v}_{i}[k],Q_{i},R_{i}\}}$)} from all the subsystems is needed to solve the problem,
which can be viewed negatively in a confidentiality point of view.

So, we use the \emph{primal decomposition} method, sometimes called \emph{quantity decomposition} or \emph{resource allocation}, with wich we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.

\subsection{Primal Decomposition based dMPC}\label{ssec:dMPC}

The technique consists of decomposing the monolithic \mpc{} in~\eqref{eq:GOP} into $M$ modified \mpc{} problems~\eqref{eq:DOP_local}, called \emph{local problems}, which can be solved in parallel by each subsystem, and a \emph{master problem}~\eqref{eq:DOP_master}, which is equivalent to the global problem and is solved by a coordinator.
\begin{subequations}
  \begin{equation}
    \left.
      \small
        J_{i}^{\star}(\thetaik)=
        \begin{matrix}
        \underset{\uiseq}{\mathrm{minimize}}&\obji\\
        \mathrm{subject~ to} &\eqref{eq:systems}\ \&\ \eqref{eq:u_local_constraints}\\
        &\Gamma_{i}\uik\preceq\qik:\dik\\
      \end{matrix}
    \right\}
    \small
    \begin{aligned}
      &\forall i\in\set{M}\\
      &\forall j\in\set{H}
    \end{aligned}
    \label{eq:DOP_local}
  \end{equation}

  \begin{equation}
    \small
    \begin{aligned}
      J^{\star}=
      \begin{matrix}
        \underset{\qiseq}{\mathrm{minimize}} &\sum_{i\in\set{M}} J^{\star}_i(\qik)\\
        \mathrm{subject~ to} &
        \left.
          \begin{aligned}
            \quad \sum_{i\in\set{M}}\qik\preceq\umax\\
            \quad \qik[k]\succeq\0_{n_{u}},       \forall i\in\set{M}
          \end{aligned}
        \right\}
        \begin{aligned}
          \forall j\in\set{H}
        \end{aligned}
      \end{matrix}
    \end{aligned}
    \label{eq:DOP_master}
  \end{equation}
\end{subequations}
The modified \mpc{} problems in~\eqref{eq:DOP_local} are created by exchanging the $\vec{u}_{\max}$ in~\eqref{eq:global_constraint} by vectors $\qik$, where which one corresponds to the quantity of the total resource $\vec{u}_{\max}$ that is allocated to agent $i$ in time $k$ for each prediction $j$, thus the names \emph{resource allocation} and \emph{quatity decomposition}. This new set of constraints have associated dual variables ${\dik\succeq\0_{n_{u}}}$. The sequence of this variables ${\qiseq}$ and ${\diseq}$ can be aggregated in vectors ${\thetaik=\kron{\1_{\predhorz}}{\qik}}$, and ${\lambdaik=\kron{\1_{\predhorz}}{\dik}}$.

The \emph{master problem} finds the optimal allocation sequences $\optthetai$ by using a projected sub-gradient method:
\begin{equation}
  \label{eq:projectedSubgradient}
\vec{\theta}[k]\pplusone=\Proj^{\set{S}}(\vec{\theta}[k]\p-\rho\p\vec{g}[k]\p),
\end{equation}
where ${\set{S} = \setbuild{\vec{\theta}}{I_{c}^{M}\vec{\theta}\preceq \vec{U}_{\max}\ \&\ \vec{\theta}\succeq\0 }}$, being ${c=\predhorz n_{u}}$, ${I_{c}^{M}=\kron{\1_{M}}{I_{c}}}$, ${\vec{U}_{\max}=\kron{\1_{\predhorz}}{\vec{u}_{\max}}}$, $(p)$ is a given step in the iterative process and $\vec{g}\p[k]$ is a sub-gradient of the objective function of problem in~\eqref{eq:DOP_master} in step $(p)$.

From~\cite{BoydEtAl2015}, it is known that $-\lambdaik$ are sub-gradients of objective $J[k]$.
Plugging it in $\vec{g}[k]\p$ we have
\begin{equation}
  \label{eq:thetaNegot}
\vec{\theta}[k]\pplusone=\Proj^{\set{S}}(\vec{\theta}[k]\p+\rho\p\vec{\lambda}[k]\p),
\end{equation}
\newcommand{\negotiation}{\emph{negotiation}}
which henceforth is referred as the \negotiation.

Once the \negotiation{} for a given time $k$ converges, the $\optuikk$ found in the last step of the \negotiation{} are applied in their corresponding agent and then the \rhs{} is followed.

The algorithm to find the optimal $\optthetai[k]$ can be summarized in algorithm~\ref{alg:quantityAlg}, and the exchange between coordinator and agents can be seen in Fig.~\ref{fig:dmpc_graph}.

\begin{algorithm2e}[h]
  \DontPrintSemicolon%
  $p:=0$\;
  Coordinator initializes $\vec{\theta}\p$ \;
  \Repeat{$\|\vec{\theta}^{(p)} -\vec{\theta}^{(p-1)}\|\leq\epsilon$}{
  Subsystems solve~\eqref{eq:DOP_local}, and send $\optlambdai (\thetai\p)$\;
  Coordinator updates allocations~\eqref{eq:thetaNegot}\;
  $p:=p+1$
}
 \caption{Quantity decomposition based \acrshort{dmpc}.}\label{alg:quantityAlg}
\end{algorithm2e}

One can observe that instead of using $\mathcal{I}_{i}$, each agent only sends $\lambdaik$.
An interpretation of $\lambdaik$ is the dissatisfaction of agent $i$ with the resource $\thetaik$ allocated for it, where ${\lambdaik=\0_{c}}$ means total satisfaction.
The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations, but if false data is injected, the \negotiation{} can be driven by a malicious agent which can take advantage of this trust to favor itself, harm others or even destabilize the \negotiation{}, as shown in~\cite{NogueiraEtAl2021}.

Our main objective is to mitigate the effects of a given configuration where agents send untrustworthy
\begin{equation*}
\lambdaicheat=\gamma_{i}(\lambdai).
\end{equation*}

\section{Towards a safe dMPC}\label{sec:TSM}

Here we focus on the local problems in~\eqref{eq:DOP_local}, and study how their structure can contribute for a safe dMPC algorithm.

\subsection{Local Problems --- Formal Analysis}\label{ssec:FA}

The \emph{local problems}~\eqref{eq:DOP_local} can be rewritten in a equivalent (same solution) constrained \qp{} form:
\todo[Correct alignment maybe]{
  \begin{subequations}\label{eq:quadratic_case}
    \begin{align}
      \label{eq:4}
      \underset{\Uik}{\mathrm{minimize}}&& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
      \mathrm{s.t.}&&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\label{eq:constraints}\\
                                        &&\Uik\succeq\0_{c},
    \end{align}
  \end{subequations}
}
where ${\Uik}$ stacks the input prediction sequences, and
${\bar{\Gamma}_{i}=\kron{I_{\predhorz}}{\Gamma_{i}}}$.
The exact values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
The approach depends only on the fact that $H_{i}:\semidefpos$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.

\no{
  We choose reference tracking just to illustrate the approach. With this objective, we have
${\vik=\wik-C_{i}\xik}$, where $\wik$ is the output reference in time $k$, which yields
\begin{equation}
\small\begin{matrix*}[l]
 H_i&=&\mathcal{D}_i^T\bar{Q}_i\mathcal{D}_i+\bar{R}_i\\
\fik&=&\mathcal{D}_i^T\bar{Q}_i(\mathcal{M}_i\xik-\Wik),
\end{matrix*}
\label{eq:matrices}
\end{equation}
where $\Wik$ stacks the setpoint prediction sequence, which is supposed to be constant during the predicted horizon ${\Wik=\kron{\1_{\predhorz}}{\vec{w}_{i}[k]}}$ (no prior information).
${\mathcal{M}_{i}}$  and ${\mathcal{D}_{i}}$ are the prediction matrices of the \mpc{}.
${\bar{Q}_{i}=\kron{I_{\predhorz}}{Q_{i}}}$, and ${\bar{R}_{i}=\kron{I_{\predhorz}}{R_{i}}}$.
}
\todo[make a better figure if time permits]{
\begin{figure}[t]
  \centering
  \begin{tikzpicture}[node distance=1.5cm]
    \tikzset{
      agent/.style={draw,circle,minimum width=1cm,},
      coord/.style={draw,circle,minimum width=1cm,very thick}
    }
    % \draw[thick,blue,rounded corners=10pt] (-1.7,1.2) rectangle (1.7,-1.7);
    % \node at (1.2,-1.3) {O};
    % \node (pi) at (-2.5,.5) {$\pi$};
    % \node (phi) at (2.5,-1.) {$\phi_{c}$};

    \node[agent] (a1) {a$_{1}$};
    \node (a2) [right of=a1]  {$\dots$};
    \node[agent] (am) [right of=a2] {a$_{M}$};
    \node[coord] (c) [above of=a2] {C};

    \draw[-latex] (a1) to [out=60,in=210] node [sloped,above] {$\vec{\lambda_{1}}[k]$} (c) ;
    \draw[-latex] (c) to [out=-120,in=30] node [sloped,below] {$\vec{\theta}_{1}[k]$} (a1);

    \draw[-latex] (am) to [out=150,in=-60]node [sloped,below] {$\vec{\lambda}_{M}[k]$} (c);
    \draw[-latex] (c) to [out=-30,in=120] node [sloped,above] {$\vec{\theta}_{M}[k]$} (am);

  \end{tikzpicture}
  \caption{Exchange between coordinator (C) and agents $a_{i}$.}\label{fig:dmpc_graph}
\end{figure}
}

We can get an explicit solution for the dual variables $\vec{\lambda}_{i}[k]$, which are \acrfull{pwa} functions w.r.t. $\vec{\theta}_{i}[k]$:
% \begin{equation}
%   \resizebox{.88\columnwidth}{!}{$
%     \begin{aligned}
%       \label{eq:lambdafuntheta}
%       \lambdaik=
%       \begin{cases}
%         -P_{i}^{1}\thetaik-\vec{s}_{i}^{1}[k]&\text{if}\ G^{1}[k]\thetaik \preceq b^{1}[k] \\
%         \qquad\quad \vdots&\qquad\quad \vdots\\
%         -P_{i}^{N}\thetaik-\vec{s}_{i}^{N}[k]&\text{if}\ G^{N}[k]\thetaik \preceq b^{N}[k] \\
%       \end{cases}
%     \end{aligned}
%     $}
% \end{equation}
\begin{equation}
\resizebox{.88\columnwidth}{!}{$
  \begin{aligned}\label{eq:lambdafuntheta}
    \lambdaik=
    % \begin{cases}
      -P_{i}^{n}\thetaik-\vec{s}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq b^{n}[k],
    % \end{cases}
  \end{aligned}
  $}
\end{equation}
with $n\in\{1\mathbin{:}N\}$.

The halfspaces defined by the set of pairs $(G^{n}[k],b^{n}[k])$ represent the combinations of active constraints~\eqref{eq:constraints} for a given time $k$, which vary w.r.t. $\vec{v}_{i}[k]$.

The $P_{i}^{n}$ are constructed using $H_{i}$ and $\bar{\Gamma}_{i}$; and $\vec{s}_{i}^{j}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$. The specific construction depends on the active constraints in~\eqref{eq:constraints}.
For example, if all constraints are active, we have ${P_{i}^{\text{ac}}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{\text{ac}}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
And if all constraints are inactive, we have ${P_{i}^{\text{in}}=0_{c\times c}}$ and ${\vec{s}_{i}^{\text{in}}=\0_{c}}$.

\begin{remark}\label{rmk:P_constant}
The $\vec{s}_{i}^{j}[k]$ depend on time $k$, while $P_{i}^{j}$ do not.
\end{remark}

\begin{challenge}\label{ch:zone_unknown}
  It is important to note that since $\vec{v}_{i}$ is unknown by the coordinator, it cannot anticipate the partition of the space for each agent.
\end{challenge}
\begin{challenge}\label{ch:s_unknown}
 For the same reason, the values of $\vec{s}_{i}[k]$ are also unknown.
\end{challenge}

 As each constraint can be active or inactive, for a given group of $t$ constraints, we can have at most $N=2^{t}$ different combinations of active sets.

 \begin{assumption}
   We assume that none of the combinations of active/inactive constraints are redundant (no linear dependency), nor make the optimization infeasible (no empty intersection).
   Thus we always have $N$ zones.
 \end{assumption}

\subsection{The Attack}\label{ssec:attack}
We suppose the malicious agent chooses a function $\gamma_{i}(\cdot)$ to use throughout a given time $k$.
\begin{assumption}
  $\gamma_{i}(\cdot)$ does not change during the \negotiation{} phase for a given time $k$.
\end{assumption}

\begin{assumption}
  The agent chooses a linear function
  \begin{equation}
\tilde{\lambdai}=\gamma_{i}(\vec{\lambda}_{i})=\Tik\vec{\lambda}_{i},
\end{equation}
  where $\determinant{T}\neq 0$.
\end{assumption}

Such attack function results in a \pwa{} solution for $\lambdaicheat[k]$:
% \begin{equation}
%   \resizebox{.88\columnwidth}{!}{$
%     \label{eq:linear_cheating}
%     \lambdaicheat[k]=
%     \begin{cases}
%       -\Tik P_{i}^{1}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{1}[k]&\text{if}\ G_{1}[k]\thetai \preceq b_{1}[k] \\
%       \qquad\quad \vdots&\qquad\quad \vdots\\
%       -\Tik P_{i}^{N}\vec{\theta}_{i}-\Tik \vec{s}_{i}^{N}[k]&\text{if}\ G_{N}[k]\thetai \preceq b_{N}[k],
%     \end{cases}
%     $}
% \end{equation}
\begin{equation}
  \resizebox{.88\columnwidth}{!}{$
    \label{eq:linear_cheating}
    \lambdaicheat[k]=
    % \begin{cases}
      -\tilde{P}_{i}^{n}[k]\thetaik-\tilde{\vec{s}}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq b^{n}[k],
    % \end{cases}
    $}
\end{equation}
with $n\in\{1\mathbin{:}N\}$, $\tilde{P}_{i}^{n}[k]=\Tik P_{i}^{n}$ and $\tilde{\vec{s}}_{i}^{n}[k]=\Tik\vec{s}_{i}^{n}[k]$.
Observe that as the function is applied only in $\lambdaik$, it does not affect the hyperplanes generating the zones.

We fix the index for these zones where all constraints are active to $1$, and we can call them the $1$-zones.

\subsection{Detection and mitigation}\label{ssec:DM}
Supposing we have a sequence of $\thetaik$ in a zone $j$, we can estimate the parameters for this zone with
\begin{equation}
  \label{eq:lambdafuntheta_tilde}
\lambdaicheat[k]=\gamma_{i}(\lambdai(\thetaik))=-\widehat{\tilde{P}_{i}^{j}}[k]\thetaik-\widehat{\tilde{\vec{s}}_{i}^{j}}[k].
\end{equation}

\begin{assumption}\label{ass:Pnominal}
  The nominal value of $P_{i}^{j}$ for this given zone $j$, denoted $\bar{P}_{i}^{j}$, is available from reliable attack-free historical data.
\end{assumption}

Since $P_{i}^{n}$ do not change w.r.t.\ time, we can detect a deviation from nominal behavior using ${E_{i}[k] =\|\widehat{\tilde{P}_{i}^{j}}[k]-\bar{P}_{i}^{j}\|_{F}}$.
Let ${d_{i}}$ be an indicator
\begin{equation}
  \label{eq:2}
  d_{i}=\indicator{E_{i}[k]\geq\epsilon_{P}},
\end{equation}
which detects the attack in agent $i$
if the disturbance $E_{i}[k]$ disrespects an arbitrary bound $\epsilon_{P}$.

If an attack is detected, we can estimate the inverse of $T_{i}(k)$ with
\begin{equation}
\widehat{{T_{i}(k)}^{-1}}=\bar{P}_{i}^{j}{\widehat{\tilde{P}_{i}^{j}}[k]}^{-1},
\end{equation}
if $\widehat{\tilde{P}_{i}^{j}}[k]$ is invertible, and $\widehat{\tilde{P}_{i}^{j}}[k]$ is only invertible when all constraints are active, that means, when ${j=1}$.
\begin{assumption}
  We suppose there exists a $1$-zone for each agent $i$ for every time $k$.
\end{assumption}
And from~\eqref{eq:lambdafuntheta}, we can derive a method to reconstruct $\vec{\lambda}_{i}$:
\begin{equation}
  \label{eq:lambda_reconstruction}
  {\vec{\lambda}_{i}}_{\mathrm{rec}}=\widehat{{T_{i}[k]}^{-1}} \tilde{\vec{\lambda}_{i}}
\end{equation}

In order to estimate $\widehat{\tilde{P}_{i}^{1}}[k]$, we must have enough observed points in the $1$-zones, so
we propose to generate points surrounding arbitrary $\bar{\vec{\theta}_{i}}$ in the $1$-zones.

\begin{assumption}\label{ass:}
  Given the unconstrained solution of~\ref{eq:quadratic_case} ${ {\optuncUik}=-H_{i}^{-1}\fik}$, we suppose
$\bar{\Gamma}_{i}\optuncUik\succeq\0_{c}$ for all $k$, that means, the points $\thetai=\0_{c}$ are in the $1$-zones.
\end{assumption}

\glsreset{EM}
Unfortunately, since we don't know the hyperplanes that separates the different zones, we don't know how close of $\bar{\theta}_{i}=\0_{c}$ we need to generate our points.
So we generate points arbitrarily close to $\bar{\theta}_{i}$ and we use the \EM{} algorithm which can potentially identify the parameters of all $N$ modes.
\subsection{Expectation Maximization}
As the method is the same for all agents and repeated each time $k$, we drop the subscript $i$ and the time dependency $[k]$ to simplify the notation.

Given a set of parameters $\set{P}$, a set of observable data $\set{B}$ and a set of non-observable data $\set{U}$, the main objective of the \EM{} algorithm is to find estimators of $\set{P}$ which maximize the log marginal likelihood of the observed data ${\ln\probability{\set{B};\set{P}}}$, for models with latent variables in $\set{U}$.

Since maximizing ${\ln\probability{\set{B};\set{P}}}$ does not have an analytical solution, the algorithm solves the optimization problem in a iterative manner.
So, rather than finding the $\set{P}$ that maximizes ${\ln\probability{\set{B};\set{P}}}$, we find the $\set{P}$ that maximizes the expectation of the complete data log-likelihood $\ln\probability{\set{B},\set{U};\set{P}}$ w.r.t.
the posterior probabilities ${\probability{\set{U}|\set{B};\set{P}}}$
calculated using a given set of parameter estimates $\set{P}_{\mathrm{cur}}$.
This provides a convergence-guaranteed iterative algorithm ensuring at each iteration a monotonic increase of the log-marginal likelihood.

The algorithm can be divided into 2 steps presented in algorithm~\ref{alg:em}.

\SetKwBlock{Estep}{ E step:}{}
\SetKwBlock{Mstep}{ M step:}{}
\begin{algorithm2e}[h]
  \DontPrintSemicolon%
  Initialize parameters $\set{P}_{\mathrm{new}}$\;
  \Repeat{$\set{P}_{\mathrm{cur}}$ converge}{
    $\set{P}_{\mathrm{cur}}=\set{P}_{\mathrm{new}}$\;
    \Estep{
      Evaluate ${\probability{\set{U}|\set{B};\set{P}_{\mathrm{cur}}}}$\;
    }
    \Mstep{
      Reestimate parameters using:
      \begin{equation}
      \resizebox{.68\columnwidth}{!}{$
        \label{eq:m_step}
        \set{P}_{\mathrm{new}}=\arg\underset{\set{P}}{\max}\ \expectation[{\probability{\set{U}|\set{B};\set{P}_{\mathrm{cur}}}}]{\ln\probability{\set{B},\set{U};\set{P}}}
        $}
      \end{equation}
    }
}
 \caption{Expectation Maximization}\label{alg:em}
\end{algorithm2e}

For a group of $O$ exchanges between an agent and the coordinator, we observe the input and response variables, identified as  ${\random{\vec{\theta}}_{o}}$ and ${\random{\vec{\lambda}}_{o}}$, ${o\in\set{O}=\{1\mathbin{:}O\}}$, which can be organized in ${\random{\Theta},\random{\Lambda}\in\R^{c\times O}}$. ${(\random{\Theta},\random{\Lambda})}$ is our set of observable data $\set{B}$.

As~\eqref{eq:linear_cheating} gives us a multidimensional \pwa{} function, we propose to use a multidimensional expansion of the model referred as \emph{mixture of linear regressions} in~\cite{FariaSoromenho2010} to map the relation between $\random{\Theta}$ and $\random{\Lambda}$. We will call the model \emph{mixture of affine regressions}, since our regressors have linear and constant terms:
% TODO(accacio): change subscript n
% \begin{equation}\label{eq:linear_cheating_random}
%   \randomvec{\lambda}_{o}=
%   \begin{cases}
%     -\tilde{P}^{1}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{1}&\text{if in zone } 1\\
%     \qquad\quad \vdots&\qquad\quad \vdots\\
%     -\tilde{P}^{Z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{Z}&\text{if in zone } Z\\
%   \end{cases}.
% \end{equation}
\begin{equation}\label{eq:linear_cheating_random}
  \randomvec{\lambda}_{o}=
    -\tilde{P}^{z}\randomvec{\theta}_{o}-\tilde{\vec{s}}^{z}\text{, if in zone } z,\\
\end{equation}
with  ${z\in\set{Z}=\{1\mathbin{:}Z\}}$.

As \eqref{eq:linear_cheating_random} is a \pwa{} function whose modes depend on the unknown zone ${z\in\set{Z}}$,
we associate to each couple $(    \randomvec{\lambda}_{o}, \randomvec{\theta}_{o})$ a latent unobserved random variable ${\random{z}_{o}}$ that indicates in which zone in $\set{Z}$ the observable variables were obtained.
These variables are organized in ${\random{Z}:\R^{1\times O}}$ which is our set of latent variables $\set{U}$.

The latent variable $\random{z}_{o}$ follows a categorical prior distribution, with associated probabilities ${\Pi=\{\pi^{z}|z\in\set{Z}\}}$ such
\[\probability{\random{z}_{o}=z}=\pi^{z} \in [0,1], \qquad \sum_{z\in\set{Z}} \pi^{z} = 1.
\]
As parameters to estimate, we have ${\set{P}=\setbuild{\set{P}^{z}}{z\in\set{Z}}}$, with ${\set{P}^{z}=(\tilde{P}^{z},\tilde{\vec{s}}^{z},\pi^{z})}$.

Since $\vec{\theta}$ is our input, we consider a non-informative improper probability density function~\citep{ChristensenEtAl2010}
\[
  \probability{\randomvec{\theta}_{o}} \,\propto \,1.
\]

Given the input and latent variables, the response variable $\random{\lambda}_{o}$ is modeled as a multivariate normal random variable with probability density function
\begin{equation}
  \label{eq:multivariate_gaussian}
\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o},\random{z}_{o}=z; \set{P}^{z}} = \mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}}),
\end{equation}
where, following~\eqref{eq:linear_cheating_random}, the mean vector is defined by
\({f(\randomvec{\theta}_{o};(P,\vec{s},\pi))=-{P}\randomvec{\theta}_{o}-{\vec{s}},}\)
and the covariance matrix ${\Sigma^{z}}$ tends to $0$.

\no{
This mixture of affine regression model corresponds to the following factorization of the complete-data (i.e., observed and latent data) likelihood, \cite{Bishop2006}:
\begin{align}\label{eq:completedataLikelihood}
  \probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}= \prod_{\set{O}}\prod_{\set{Z}}\big[&\probability{\randomvec{\lambda}_{o}|\randomvec{\theta}_{o},\random{z}_{o}=z;\set{P}^{z}} \nonumber \\
& \times \probability{\random{z}_{o}=z}\probability{\randomvec{\theta}_{o}}\big]^{\indicator{\random{z}_{o}=z}},
\end{align}
}

We can calculate the posterior probabilities  $\zeta_{zo}(\set{P})=\probability{\random{z}_{o}=z|\randomvec{\lambda}_{o},\randomvec{\theta}_{o};\set{P}}$, also called \emph{responsibilities}:
\begin{align}
  \label{eq:responsibilities}
\zeta_{zo}(\set{P})&=\frac{\pi_{z}{\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}})}}{\sum\limits_{j=1}^{Z}\pi_{j}\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{j}),{\Sigma^{j}})},
\end{align}
and then calculate the expectation of ${\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}}$ with respect to ${\zeta_{zo}(\set{P}_{\mathrm{cur}})}$ \citep[Chapter 9]{Bishop2006}
\begin{align}
  \label{eq:completedataLogLikelihood_expectation}
\expectation[{\zeta_{zo}(\set{P}_{\mathrm{cur}})}]{\ln\probability{\random{\Theta},\random{\Lambda},\random{Z};\set{P}}}&= \sum_{o\in\set{O}}\sum_{z\in\set{Z}}  \zeta_{zo}(\set{P}_{\mathrm{cur}})\alpha_{zo},
\end{align}
where ${\alpha_{zo}=\ln{\pi_{z}}+\ln{\mathcal{N}(\randomvec{\lambda}_{o};f(\randomvec{\theta}_{o};\set{P}^{z}),{\Sigma^{z}})}}$.

\begin{remark}
The $\Sigma^{z}$ used in~\eqref{eq:responsibilities} are updated using a technique called \emph{Simulated annealing}~\citep{OzerovFevotte2010}, where they are initialized with arbitrarily big values indicating the uncertainty of the parameters and it is reduced as the estimations converge.
\end{remark}

Here we introduce a variable ${\vec{\phi}^{z}={[\vectorize{\tilde{P}^{z}}^{T}\ {(\tilde{\vec{s}}^{z})}^{T} ]}^{T}}$.
Using the KKT conditions, we can find an optimal solution for the problem in~\eqref{eq:m_step}, by
taking the gradients of~\eqref{eq:completedataLogLikelihood_expectation} with respect to vectors $\vec{\phi}^{z}$ and making them vanish.
Because of the multidimensional nature of the problem, some matrix operations are needed to synthesize the results.
After those operations, we have a matricial solution that yields the optimal estimates $\vec{\phi}^{z}_{\mathrm{new}}$:
\begin{equation}
  \label{eq:mstepestimation}
  \vec{\phi}^{z}_{\mathrm{new}}=\pseudoinv{(\Xi^{z}\random{\Omega})}\Xi^{z}\vectorize{\random{\Lambda}},
\end{equation}
where
${\random{\Omega}=[\hadamard{(\Upsilon \random{\Theta}\Delta)}{Y};G]}$,
with matrices
${\Upsilon=\kron{\1_{c}^{T}}{I_{c}}}$,
${\Delta=\kron{I_{O}}{\1_{c}^{T}}}$,
${Y=\kron{G}{\1_{c}}}$,
${G=\kron{\1_{O}^{T}}{I_{c}}}$,
and
\[{\Xi^{z}={\diag(\sqrt{{\zeta(z_{z1};\set{P}_{\mathrm{cur}})}}I_{c},\cdots,\sqrt{{\zeta(z_{zO};\set{P}_{\mathrm{cur}})}}I_{c})}}.\]
Doing the same for $\pi^{z}$ we get
\begin{equation}
  \label{eq:3}
  \pi^{z}=\sum\limits_{o\in\set{O}}\tfrac{\zeta_{zo}(\set{P}_{\mathrm{cur}})}{O}.
\end{equation}

As we can see,~\eqref{eq:mstepestimation} is the solution of a weighted Least-Squares, having the responsibilities as weights, adjusting the contribution of all observations to the regression model of index $z$.
We can see some similarities to the K-planes algorithm (see~\cite{BradleyMangasarian2000}), but \EM{} is more compromising.
Instead of affecting the observed data to a zone with 100\% of certainty (\emph{hard assignment}), \EM{} uses the responsibilities (\emph{soft assignment}) for each zone.

Once the estimates $\vec{\phi}_{z}^{\mathrm{new}}$ converge, we can reconstruct the estimates $\tilde{P}^{z}$ and $\tilde{\vec{s}}^{z}$, and use in our mitigation scheme proposed in \S\ref{ssec:DM}.
Observe that not necessarily the $z$ indices correspond to the indices $n$ in~\eqref{eq:lambdafuntheta}.
So, to find the $z$ that corresponds to ${n=1}$, we take the observation $o$ for $\bar{\theta}=\0_{c}$, which we know belongs to the $1$-zone, and we find the most probable $z$ with
\begin{equation}\label{eq:argmaxz}
\arg\underset{z}{\max}\ {\zeta_{zo}(\set{P})}.
\end{equation}

\begin{remark}
A discussion about the initialization and update of other parameters is beyond the scope of this article, so we refer the reader to~\cite{Bishop2006}, and~\cite{DempsterEtAl1977}.
\end{remark}

\subsection{Safe dMPC Algorithm}\label{ssec:safe_algo}
Integrating the EM to our detection and mitigation mechanism we can propose in Algorithm~\ref{alg:safeDMPC} a secure \dmpc{}, divided into two phases: detection and negotiation.
\SetKwBlock{negotPhase}{ Negotiation Phase:}{}
\SetKwBlock{detectPhase}{ Detection Phase:}{}
\begin{algorithm2e}[h]
  \DontPrintSemicolon
  \detectPhase{
    Coordinator sends sequence of $\vec{\theta}_{i}^{o}$, $o\in\set{O}$ \;
    Subsystems solve~\eqref{eq:DOP_local}, and send $\tilde{\vec{\lambda}}^{o}_{i}$, $o\in\set{O}$\;
    Coordinator estimates $\widehat{\tilde{P}_{i}^{1}}[k]$ and $\widehat{\tilde{\vec{s}}_{i}^{1}}[k]$ with \EM{}\;
    Coordinator computes $d_{i}$ using~\eqref{eq:2}\;
  }
  \negotPhase{
  Apply Algorithm~\ref{alg:quantityAlg} with adequate versions of $\vec{\lambda}_{i}\p$:\;
  \quad$\tilde{\vec{\lambda}}_{i}\p$, if $d_{i}=0$ and ${\vec{\lambda}_{i}}_{\mathrm{rec}}$, if ${d_{i}=1}$~\eqref{eq:lambda_reconstruction}\;
 }
 \caption{Secure DMPC.}\label{alg:safeDMPC}
\end{algorithm2e}

\section{Example: Temperature Control}\label{sec:App_dMPC}
The system consists of $4$~distinct~rooms (I to IV), which we want to control the air temperatures inside each one of them.
The systems are modeled as continuous-time linear \mbox{time-invariant} systems using the \mbox{3R-2C} model~\citep{NogueiraEtAl2021}, with parameters shown in Tables~\ref{tab:modelParamMeaning} and~\ref{tab:modelParam}, and state space matrices
\begin{equation*}
  \label{eq:room_system}
  \begin{matrix}
\resizebox{.98\columnwidth}{!}{$
  A_{\mathrm{c}_{i}}=\left[
    \begin{matrix}
      -\frac{1}{C^{\text{walls}}_{i}R^{\text{oa/ia}}_{i}}-\frac{1}{C^{\text{walls}}_{i}R^{\text{iw/ia}}_{i}}& \frac{1}{C^{\text{walls}}_{i}R^{\text{iw/ia}}_{i}}\\
      \frac{1}{C^{\text{air}}_{i}R^{\text{iw/ia}}_{i}} &-\frac{1}{C^{\text{air}}_{i}Ro_{i}}-\frac{1}{C^{\text{air}}_{i}R^{\text{iw/ia}}_{i}}
    \end{matrix}\right]
    $}
  \\
  \begin{matrix}
    B_{\mathrm{c}_{i}}=\left[
      \begin{matrix}  \frac{10}{C^{\text{walls}}_{i}}& 0\end{matrix}
    \right]\T&C_{\mathrm{c}_{i}}=\left[\begin{matrix}1 & 0\end{matrix}\right]
  \end{matrix}
  \end{matrix},
\end{equation*}
where the states are ${\vec{x}_{i}=[{{x}_{A}}_{i}\T\ {{x}_{W}}_{i}\T]\T}$, which represent the mean temperatures of the air and walls inside room $i$, and the input of the system is $\vec{u}_{i}$, which is the heating power
for the corresponding room.

The inputs are constraint by ${\sum_{i=1}^{4}\vec{u}_{i}(k)\preceq 4\mathrm{kW}}$.

\begin{table}[b]
  \centering
  \caption{Model Parameters}\label{tab:modelParamMeaning}
  \begin{tabular}[b]{cl}
    \toprule
    Symbol&Meaning\\
    \midrule
    $C^{\text{air}}_{i}$&Heat Capacity of Inside Air\\
    $C^{\text{walls}}_{i}$&Heat Capacity of External Walls\\
    $R^{\text{oa/ia}}_{i}$&Resistance Between Inside and Outside Air (from windows)\\
    $R^{\text{iw/ia}}_{i}$&Resistance Between Inside Air and Inside Walls\\
    $R^{\text{ow/oa}}_{i}$&Resistance Between Outside Air and Outside Walls\\
    \bottomrule
  \end{tabular}
\end{table}

  \begin{table}[b]
  \centering
  \caption{
\todo{
  change values
}
    Value of parameters for each agent}\label{tab:modelParam}
  \begin{tabular}[t]{cccccc} \toprule
    Symbol& I & II & III & IV &Unit\\
    \midrule
    $C^{\text{walls}}$    &$5$   & $4$ & $4.5$ &$4.7$ &$10^{4}\mathrm{J/K}$ \\
    $C^{\text{air}}$               &$8$   & $7$ &$9$&$6$  &$10^{4}\mathrm{J/K}$  \\
    $R^{\text{oa/ia}}$               &$5$   & $6$& $4$& $5$  &$10^{-3}\mathrm{K/W}$ \\
    $R^{\text{iw/ia}}$               &$2.5$ & $2.3$&$2$&$2.2$ &$10^{-4}\mathrm{K/W}$\\
    $R^{\text{ow/oa}}$               &$0.5$ & $1$ &$0.8$&$0.9$  & $10^{-4}\mathrm{K/W}$ \\
    \bottomrule
  \end{tabular}
\end{table}

The subsystems are discretized using the zero-order hold discretization method with sampling time
${T_{s}=0.25\mathrm{h}}$
and the quantity decomposition-based DMPC is implemented using prediction horizon ${\predhorz=4}$.

Three scenarios are simulated for a period of 12.5 hours:

  \begin{enumerate}
    \item Nominal behavior.
    \item \mbox{Agent I presents non-cooperative behavior} ${T_{I}[k]}$ for ${k\geq25}$, without correction.
    \item \mbox{Agent I presents non-cooperative behavior} ${T_{I}[k]}$ for ${k\geq25}$, with correction, ${\epsilon_{P}=10^{-4}}$.
  \end{enumerate}

In Fig.~\ref{fig:response3Scenarios}, first, we compare the air temperature in room I
with its reference (${w_{I}=25.5^{\circ}}$C), and then the
decision variable ${E_{I}(k)}$ with the threshold $\epsilon_{P}$.
Observe that in the nominal behavior (in magenta), the reference $w_{I}$ is not reached
due to power constraints the systems are under.

\begin{figure}[h]
  \centering
 \includegraphics[width=\columnwidth]{airtemp_roomI/__ErrorWX_command_normErrH.pdf}
  \caption{Air temperature in room I and the decision variable $E_{I}[k]$ for different scenarios: nominal (N), selfish behavior without correction (S),
and selfish behavior with correction.}\label{fig:response3Scenarios}
\end{figure}

As expected, the decision variable lies under the threshold $\epsilon_{P}=10^{-4}$ with values of order ${E_{I}^{N}(k)\approx10^{-10}}$.

When the agent presents a selfish behavior (in orange), the tracking error
${w_{I}-y_{I}}$ is reduced but insufficient to attain the reference.
In this case, the detection variable surpasses $\epsilon_{P}$,
${E_{I}^{S}\approx200}$, indicating the change of behavior.

When the correction is activated in the system, we see that the temperatures approach their nominal value $y_{I}^{N}$.
We can also illustrating the good
performances of our proposition by comparing the inputs in Fig.~\ref{fig:control_3Scenarios}. When I is selfish its control increases while the other decreases, unless the correction is activated and it approaches the nominal.

We can also evaluate the performance of the proposed mechanism by comparing the global objective function calculated using the total period of simulation ${N=50}$ for the 3 scenarios (Table~\ref{tab:costsGlobalLocal}).
When agent I is selfish, we see the increase in the objective, degrading the overall performance.
When the correction mechanism is activated, the difference between costs is minimal, $|\frac{J^{C}-J^{N}}{J^{N}}|\leq10^{-8}$.
%%%%%
% TODO(accacio): Verify
\begin{table}[t]
  \centering
  \caption{Comparison of costs $J_{i}$}\label{tab:costsGlobalLocal}
  \begin{tabular}[t]{cccc} \toprule
    Agent  & Nominal & Selfish & Selfish + Correction\\
    \midrule
    % I & -11468 &  -9074 & -11468 \\
    % II & -6536 &  -7151 & -6536 \\
    % III & -3192 &  -4052 & -3192 \\
    % IV & -2849 &  -3694 & -2849 \\
    % Global & -24044 &  -23971 & -24044 \\
    Global & -24044.180111 &  -23971.164122 & -24044.180027\\
    \bottomrule
  \end{tabular}
\end{table}
% \enlargethispage{-.8cm}

\section{CONCLUSION AND FUTURE WORKS}\label{sec:CC}
\todo{
In this paper, an algorithm for monitoring and correcting exchanges between agents in a resource-sharing system has been proposed.
The algorithm exploits the particular structure of exchanges, part of which must be constant over time.
The first phase consists of identifying this constant part and checking if an attacker has modified it.
From this identification, it is possible to reconstruct the original mechanism and find the centralized optimality.
This principle should be generalized to other types of decomposition structures, and this is what we plan to do in the near future.
}

\begin{figure}[h]
  \centering
 \includegraphics[width=\columnwidth]{airtemp_roomI/control.pdf}
  \caption{Control applied in all rooms for the 3 scenarios.}\label{fig:control_3Scenarios}
\end{figure}

% \nocite{*}
\bibliography{bibliography}
% \bibliographystyle{ifacconf-harvard}
% \todo[Change bibliography to the one used in the aux]{}
\end{document}